---
title: "p039_normative_long_post_save_reliability_data"
author: "DVP"
date: "2022.12.12"
output: html_document
---

# copute reliability for aseg + destrieux ROIs 
save within subject change in the slopes from multi cohort dataset
save error from s2c

# specify folders and files
## setup
```{r setupp}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rio)

# set directories
options(bitmapType = "cairo")
outdir=here('data_normative_long/df_mri/all')
s2dir = here("data_flair/s2c")
reliabilitydir = file.path(outdir, "reliability")
errordir = file.path(reliabilitydir, "error")
wsdir = file.path(reliabilitydir, "ws")
scriptsdir = here("scripts")
```

## set name to common files
```{r common files}
df.harmonize = import(file.path(outdir, 'harmonize_variables.xlsx'))
sex.rename = import(file.path(outdir, 'harmonize_variables.xlsx'), sheet = "sex")
phenotypes.s2c = phenotypes.gamm = df.harmonize$normative_modelling[7:dim(df.harmonize)[1]]
phenotypes.s2c[grepl("Thalamus", phenotypes.s2c)] = c("Left-Thalamus" , "Right-Thalamus")
phenotypes.s2c = gsub("&", "_and_", phenotypes.s2c)

load(file.path(outdir, "df.all.filt.Rda"))
model = "reliability/ws"
```

# get ws from multi-chohort dataset. 
## extract ws from data
```{r compute normative scores}
scriptname =  here("scripts/normative_scripts/submit_compute_gamm_Zchange.sh")

for (i in 1:length(phenotypes.gamm)) {
  system(paste(
    "module purge;",
    "sbatch",
     scriptname, 
     outdir, 
     model,
     paste0("'",phenotypes.gamm[i],"'"),
     sep = " "))
}
```

## merge ws computational parameters from different regions
```{r computational parameters}
model_parameters = lapply(phenotypes.gamm, 
                          function(x) {import(file.path(outdir, model, x,"model_parameters.csv"))})
model_parameters = reduce(model_parameters, cbind) %>% 
  t() %>% 
  data.frame()
names(model_parameters) = 
  c("mean60",
    "mean65",
    "mean70",
    "mean76",
    "mean80",
    "sd", 
    "pct_60.80",
    "pct_65.80",
    "pct_70.80",
    "fderiv_60.80",
    "fderiv_70.80",
    "var1", 
    "var_2_2",
    "var_4_3")
model_parameters$features = phenotypes.gamm

save(model_parameters, 
          file = file.path(reliabilitydir, "full.sample.rda"))

```


# get error from s2c 
## define functions s2c
```{r define function desterieux}
fetch_aparc_data_s2c = function(idir) {
lfiles = list.files(idir, pattern = ".txt")
subs = 
  strsplit(lfiles,"sub-") %>% 
  reduce(rbind) %>% 
  .[,2] %>% 
  substr(., 1,7) %>% 
  unique()

df.mri = lapply(lfiles, function(x) { import(file.path(idir, x))})
for (df in 1:length(df.mri)) {
  names(df.mri[[df]])[1] <- "input"
}
names(df.mri) = lfiles

df.mri.all = list()
for (i in 1:length(subs)) {
  idx = lfiles[grep(subs[i], lfiles)]
  x = df.mri[idx]
  df.mri.all[[i]] = x %>% reduce(cbind)
}
names(df.mri.all) = subs
df.mri = data.table::rbindlist(df.mri.all, idcol = "subs")
return(df.mri)
}

work_with_residuals = function(db) {
  x = 
      db %>% 
        group_by(subs) %>% 
        mutate(n = n_distinct(input)) %>% 
        filter(n > 1)  %>% 
    pivot_longer(-c(subs, input), 
                      names_to = "features", 
                      values_to = "values") %>% 
    mutate(sess = gsub("ses-0", "", input), 
           sess = gsub(".long.base", "", sess),
           sess = as.numeric(sess)) %>% 
    arrange(subs,sess)
  
  
  mod =
  x %>% 
    group_by(subs, features) %>% 
    mutate(ref = first(values),
           sesref = first(sess),
           errM = values - ref) %>% 
    filter(!sesref == sess) %>% 
    mutate(pct.err = 100* errM / ref)
  
  
  mod = mod %>%
  mutate(pctM = median(abs(pct.err), na.rm = T),
         pctmad = mad(abs(pct.err), na.rm = T), 
         abspcterr = if_else(abs(pct.err) > pctM +  8*pctmad,NaN,abs(pct.err)))
  
  mod.sum.RID = 
    mod %>% 
    group_by(subs, features) %>% 
    summarise(pct.err.mean = mean(abspcterr, na.rm = T),
              meanV =  mean(ref, na.rm = T))
  
  mod.sum = 
  mod %>% group_by(features) %>% 
    summarise(pct.err.mean = mean(abspcterr, na.rm = T),
              meanV = mean(ref), 
              sdV = sd(ref))
  
  df.out = list()
  df.out$mod = mod
  df.out$mod.sum.rid = mod.sum.RID
  df.out$mod.sum = mod.sum
  return(df.out)
}
```


## main routine for s2c
```{r fetch data, echo=FALSE}
df = list()
df$df = fetch_aparc_data_s2c(here("data-raw/tabulated/s2c/mri_fs7_noflair"))
df$df = df$df %>% 
  select(unique(colnames(.)))
df$mod = work_with_residuals(df$df)


df.out = df$mod$mod.sum
idx = match(phenotypes.s2c,df.out$features)
df.out = df.out[idx, ]


save(
  df.out,
  file = file.path(errordir, "s2c_data.rda"))
```


# Compute error variance
## open data and merge
```{r}
# error
load(file.path(errordir, "s2c_data.rda"))
df.error = df.out


#ws variance 
load(file.path(reliabilitydir, "full.sample.rda"))
df.ws = 
  model_parameters %>% 
    dplyr::select(features, 
            var_4_3) %>% 
  rename("seD" = "var_4_3")
df.ws$features = phenotypes.s2c

load(here("data_memory_long/all/brain_variance.rda"))
df.variance = 
  var_mri %>% 
  pivot_longer(-dataset, 
               names_to = "features", 
               values_to = "variance")

load(here("data_memory_long/all/brain_overview.rda"))
df.dataset = 
  y %>% 
  rename(c("tp" = "mean_tp", 
         "fu" = "mean_followup"))

# merge datasets
df = inner_join(df.error, df.ws)
df = df %>%  filter(!grepl("_area", features))
df = df %>% mutate(features = gsub("_thickness", "", features),
                   features = gsub("Thalamus", "Thalamus-Proper", features),
              features = gsub("MeanThickness", "MeanThickness_thickness", features),
              features = gsub("3rd-Ventricle", "x3rd-ventricle", features),
              features = gsub("4th-Ventricle", "x4th-ventricle", features),
              features = gsub("Right-Cerebellum", "xRight-Cerebellum", features))


df.variance = inner_join(df.variance, 
                  df.dataset)
df = inner_join(df, df.variance)
save(df, 
     file = file.path(reliabilitydir, "reliability_parameters.rda"))
```


## get icc and error variance based on formula
  ws within-subject variability
  # for equispaced timepoints
   #e² / (t²*n*(n-1)/12*(n-1)
   # e = error
   # t = time
   # n = number of timepoints

  # bs = var(bs) = sd² = seD²
  ICC(2,1) = bs² / bs² + ws^2 
  in the context of long trajectories 
  
  Alternatively, if we did not know the error variance, but had reliability, 
  we would do it this way:
  reliability <- .8 # (I know that in this example: 1 / (1 + .5^2))
  error_variance <- (1 - reliability) * var(w)
  with this approach, the error variance is an estimate, and will change with the
  data, but that's how it always is, except when I simulate and know the truth.
  thus, the error_variance computed here will vary around its true value of 0.5^2,
  if we change the random number seed above
  
## Save reliability parameters- subcortical thickness area  
```{r}
load(file.path(reliabilitydir, "reliability_parameters.rda"))

df = 
df %>% 
  mutate(bs = seD^2, 
         ws = pct.err.mean^2*((fu^2*tp*(tp+1))/ (12*(tp-1)))^-1, 
         icc = bs /(ws + bs))

df = 
  df %>% 
  mutate(error_variance = (1 - icc) * variance)

df.features = 
  df %>% group_by(features) %>% summarise(pct.err.mean = mean(pct.err.mean), seD = mean(seD))
save(df,
     df.features,
     file = file.path(reliabilitydir, "reliability_parameters.rda"))
```

