---
title: "p039_normative_long_prepare_df"
author: "DVP"
date: "2022.12.12"
output: html_document
---

# Fetch neuroimaging data, and link with main data.files
### This script will populate rawdatafiles with 1) csv file linking available mri data with existing tables; 2) will import fs derivatives of interest in the folder
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rio)

#here()
source(here("scripts/normative_scripts/helper_normative_long_prepare_df.r"))
mrifiles = c("lh.aparc.a2009s.area.txt",
             "lh.aparc.a2009s.thickness.txt",
             "lh.aparc.a2009s.volume.txt",
             "rh.aparc.a2009s.area.txt",
             "rh.aparc.a2009s.thickness.txt",
             "rh.aparc.a2009s.volume.txt",
             "aseg.txt")
```

## This R Markdown will fetch neuroimaging data and neuroimaging summary tables and save them in the site folder. 
### Fetch neuroimaging data for uio.
```{r LCBC}
## set up links
set_links("uio")

# link tab data with fs images
check.uio.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for mpib
```{r MPIB}
## set up links  
set_links("mpib")

# link to mpib data
check.mpib.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for umu
```{r Umu_Betula}
## set up links  
set_links("umu")

# link to umu data
check.umu.fs(sitedir, outdir, mrifiles)

# 78 cogn. data missing
# 61 mri missing 
```

### Fetch neuroimaging data for ub.
```{r UB}
## set up links  
set_links("ub")

# link to umu data
check.ub.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for habs.
```{r habs}
## set up links  
set_links("habs")

# link to habs data
check.habs.fs(sitedir, outdir, mrifiles)
```

### Fetch neuroimaging data for cognorm / ous
```{r cognorm / ous}
set_links("ous")

# link to uos data
check.ous.fs(sitedir, outdir, mrifiles)
```

### Fetch neuroimaging data for aibl.
```{r aibl}
set_links("aibl")

# link to uio data
check.aibl.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for adni.
```{r adni}
set_links("adni")

# link to uio data
check.adni.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for preventAD.
```{r preventAD}
set_links("preventad")

# link to preventad data data
check.preventad.fs(sitedir, outdir, mrifiles)
```

### Fetch neuroimaging data for oasis3.
```{r oasis3}
set_links("oasis3")

# link to oasis data
check.oasis.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for bbhi.
```{r bbhi}
set_links("bbhi")

# link to bbhi data
check.bbhi.fs(sitedir, outdir, mrifiles)
```


### Fetch neuroimaging data for ukb.
```{r ukb}
set_links("ukb")

# link to uio data
check.ukb.fs(sitedir, outdir, mrifiles)
```



### Fetch neuroimaging data for vetsa
```{r vetsa}
set_links("vetsa")

# link to wayne data
check.vetsa.fs(sitedir, outdir, mrifiles)
```
