---
title: "extract_data"
author: "MS- DVP"
date: "19/04/2022"
output: html_document
---
  
# Fetch memory data, and link with main data.files
### This script will populate rawdatafiles with 1) csv file linking available mri data with existing tables; 2) will import fs derivatives of interest in the folder
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rio)
#library(readr)

#here()
source(here("scripts/memory_long_scripts/1_extract_functions.R"))


```

## This R Markdown will fetch memory data. 
### Fetch memory data for uio.
```{r LCBC}
## set up links
set_links("uio")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.uio(sitedir, outdir, mrifiles)


```

### Fetch memory data for umu
```{r Umu_Betula}
## set up links  
set_links("umu")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.umu(sitedir, outdir, mrifiles)


```

### Fetch memory data for ub.
```{r UB}
## set up links  
set_links("ub")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.ub(sitedir, outdir, mrifiles)
```

### Fetch memory data for mpib 
```{r MPIB}
## set up links  
set_links("mpib")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.mpib(sitedir, outdir, mrifiles)
```

### Fetch memory data for adni.
```{r adni}
set_links("adni")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.adni(sitedir, outdir, mrifiles)
```

### Fetch memory data for aibl.
```{r aibl}
set_links("aibl")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.aibl(sitedir, outdir, mrifiles)
```


### Fetch memory data for habs.
```{r habs}
## set up links  
set_links("habs")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.habs(sitedir, outdir, mrifiles)
```

### Fetch memory data for cognorm / ous
```{r cognorm / ous}
set_links("ous")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.ous(sitedir, outdir, mrifiles)
```


### Fetch memory data for preventAD.
```{r preventAD}
set_links("preventad")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.preventad(sitedir, outdir, mrifiles)
```


### Fetch memory data for ukb.
```{r ukb}
set_links("ukb")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.ukb(sitedir, outdir)
```

### Fetch memory data for oasis.
```{r ukb}
set_links("oasis3")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.oasis3(sitedir, outdir, mrifiles)
```

### Fetch memory data for BBHI.
```{r bbhi}
## set up links  
set_links("bbhi")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.bbhi(sitedir, outdir, mrifiles)
```

### Fetch memory data for vetsa.
```{r vetsa}
## set up links  
set_links("vetsa")

# extract memory data (UPDATE: also non longitudinal data)
extract.longdata.vetsa(sitedir, outdir, mrifiles)
```
