---
title: "postproc_MetaMega"
author: "MS"
date: '2023-08-30'
output: html_document
---

#LOAD LIB

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(sn)

options(bitmapType='cairo')

wd = here("data_memory_long/all")
mridir=here('data_normative_long/df_mri/all')
reliabilitydir = file.path(mridir, "reliability")
source(here("scripts/memory_long_scripts/6_analyses_functions.R"))
```

# 0. Prepare data

```{r merge data}
# fetch brain data and merged it with basic sociodemographics
df.mri.change = wrapper_merge_mri_data(wd)
merged = wrapper_merge_mem_with_mri_data(wd)
merged = merge_apoe(wd)
df.merged = wrapper_merge_reliability (wd, reliabilitydir) # merged.reliability.rda (main data.frame of reference)
```

# 1. MEGA - slope

## sbatch main and age interaction

```{r}
scriptname = here("scripts/memory_long_scripts/sbatch_scripts/submit_run_mega.sh")
wd = here("data_memory_long/all")
load(file.path(wd, "merged_reliability.rda"))
df.merged = remove.vars(df.merged)
feature = unique(df.merged$features)

for (i in 1:length(feature)) {
  #i = 1
   system(paste(
      "module purge;",
      "sbatch",
       scriptname, 
       feature[i],
       wd,
       sep = " ")) 
}
```

## Merge regions

```{r}
temp_folder = here("data_memory_long/all/mega/temp_models")
mylist = list.files(temp_folder, pattern = ".Rda")
mylist = mylist[!grepl("^models_", mylist)]
grot = list()
for (i in 1:length(mylist)) {
  fileee <- file.path(temp_folder, mylist[i])
  load(fileee)
  grot[[i]] = xx
}

df.mega.output = data.table::rbindlist(grot)

p.correct = c("lme.w_pval",
              "lme.slopes.w_pval", 
              "gamm.w.pval", 
              "gamm.w.pval.bootstrap",
              "gamm.int.w.pval",
              "gamm.int.w.pval.bootstrap")

p.direction = 
  tribble(~p.value, ~estimate, 
        "lme.w_pval", "lme.w_beta",
        "lme.slopes.w_pval", "lme.slopes.w_beta")

vars_sign = c("lme_beta", 
              "lme.w_beta",
              "lme.w_beta.ctrlicv",
              "lme.w_beta.ctrlicvage",
              "lme.slopes.w_beta")

mega = prepare_df2(df.mega.output, p.correct, p.direction, vars_sign)

vars.infinite =  names(mega)[grep("log", names(mega))]
mega = change_infinite_values(mega, vars.infinite, 5000)

save(mega,
       file = file.path(wd,"mega", "df.mega.rda"))
```

## make plots for gamm models

```{r, warning=F, echo =F}
mega_plots_and_derivatives(wd)
```

## get estimates and derivatives

```{r}
# load mega data
load(file.path(wd,"mega", "df.mega.rda"))
dt.predictions = wrapper_get_predictions_derivative(wd) 
dt.predictions = data.table::rbindlist(dt.predictions)
mega = left_join(mega, dt.predictions)    
mega = wrapper_get_estimates(mega)  
dt.estimates = wrapper_draw_smooth(mega)
dt.estimates = data.table::rbindlist(dt.estimates)
mega = left_join(mega, dt.estimates)    
save(mega, file =file.path(wd,"mega", "df.mega.rda"))
```

# 2. Hemispheric differences (mega)

```{r}
hemi.diff = wrapper_hemispheric_differences(wd)
```

# 3. Global or regional (mega)

## Consensus Clustering + PCA

```{r}
m3c = wrapper_clustering(wd)
pca_result = wrapper_pca(wd)

## extra analysis
scriptname = here("scripts/memory_long_scripts/sbatch_scripts/submit_run_consensuscluster_extra.sh")
types = c("all", "pca", "cl1", "basic")
for (i in 1:length(types)) {
  #i = 1
   system(paste(
      "module purge;",
      "sbatch",
       scriptname, 
       types[i],
       wd,
       sep = " ")) 
}

df.basic =retrieve_pboot_cc(wd, type = "basic")
df.pca =retrieve_pboot_cc(wd, type = "pca")
df.cl1 =retrieve_pboot_cc(wd, type = "cl1")
df.clall =retrieve_pboot_cc(wd, type = "all")
dt.basic1 = wrapper_get_weights_cc_analysis(wd, type = "basic")
dt.pca1 = wrapper_get_weights_cc_analysis(wd, type = "pca")
dt.cl11 = wrapper_get_weights_cc_analysis(wd, type = "cl1")
dt.clall1 = wrapper_get_weights_cc_analysis(wd, type = "all")

save(df.basic, 
     df.pca, 
     df.cl1, 
     df.clall, 
     dt.basic1, 
     dt.pca1, 
     dt.cl11, 
     dt.clall1, 
     file = file.path(wd,"mega","consensus_cluster","mod.results.rda"))
```

# 4. APOE (mega)

## sbatch apoe memory only and brain only

```{r}
wd = here("data_memory_long/all")
# run memory - apoe 
scriptname = here("scripts/memory_long_scripts/sbatch_scripts/submit_run_apoe_memory.sh")
system(paste(
      "module purge;",
      "sbatch",
       scriptname, 
       file.path(wd),
       sep = " ")) 

# run brain - apoe
scriptname = here("scripts/memory_long_scripts/sbatch_scripts/submit_run_apoe_brain.sh")
load(file.path(wd, "merged_reliability.rda"))
df.merged = remove.vars(df.merged)
feature = unique(df.merged$features)

for (i in 1:length(feature)) {
   system(paste(
      "module purge;",
      "sbatch",
       scriptname, 
       feature[i],
       wd,
       sep = " ")) 
}
```

## sbatch apoe - associations

```{r }
scriptname = here("scripts/memory_long_scripts/sbatch_scripts/submit_run_apoe.sh")
wd = here("data_memory_long/all")
load(file.path(wd, "merged_reliability.rda"))

df.merged = remove.vars(df.merged)

feature = unique(df.merged$features)


for (i in 1:length(feature)) {
   system(paste(
      "module purge;",
      "sbatch",
       scriptname, 
       feature[i],
       file.path(wd),
       sep = " ")) 
}
```

# 4.1 APOE BRAIN And MEmory only

## merge regions apoe on brain

```{r}
wd = here('data_memory_long', 'all')
temp_folder = here("data_memory_long/all/mega/apoe/output/temp_models_mem")
mylist = list.files(temp_folder, pattern = ".Rda")
mylist = mylist[!mylist %in% c("models.Rda", "output.Rda")]
mylist = mylist[!grepl("^models.", mylist)]
grot = list()
for (i in 1:length(mylist)) {
  fileee <- file.path(temp_folder, mylist[i])
  f = gsub(".Rda", "",mylist[i])
  load(fileee)
  output$features = f
  grot[[i]] = output
  
}
df.mega.apoe.brain = data.table::rbindlist(grot, fill=T)

p.correct = c("lme.apoe.brain_pval",
              "gam.apoe.brain_agep",
              "gam.apoe.brain_agep.boot")
 
p.direction =
  tribble(~p.value, ~estimate,
        "lme.apoe.brain_pval", "lme.apoe.brain_beta")

mega.apoe.brain = prepare_df2(df.mega.apoe.brain, p.correct, p.direction, vars_sign=c())

save(mega.apoe.brain,
       file = file.path(wd,"mega","apoe", "df.mega.apoe.brain.rda"))

#age_plot_apoe_brain()
```

# 4.2 APOE -BRAIN memorya ssociations

## Merge regions (apoe)

```{r}
wd = here('data_memory_long', 'all')
temp_folder = here("data_memory_long/all/mega/apoe/output/temp_models")
mylist = list.files(temp_folder, pattern = ".Rda")
mylist = mylist[!grepl("^models_", mylist)]
grot = list()
for (i in 1:length(mylist)) {
  fileee <- file.path(temp_folder, mylist[i])
  load(fileee)
  grot[[i]] = output
}
df.mega.apoe = data.table::rbindlist(grot, fill=T)

p.correct = c("lme.apoe.memory_pval",
              "lme.apoe.brain_pval",
              "lme.apoe_pval", 
              "gam.int.w.apoe0p", 
              "gam.int.w.apoe1p", 
              "gam.main.w.apoeOp", 
              "gam.main.w.apoe1p",
              "gam.int.w.apoeOp", 
              "gam.main.w.apoeOp",
              "gam.int.w.apoeOp.bootstrap",
              "gam.main.w.apoeOp.bootstrap",
              "gam.main.w.apoe0p.bootstrap",
              "gam.main.w.apoe1p.bootstrap",
              "gam.int.w.apoe0p.bootstrap",
              "gam.int.w.apoe1p.bootstrap")
 
p.direction =
  tribble(~p.value, ~estimate,
        "lme.apoe.memory_pval", "lme.apoe.memory_beta",
        "lme.apoe.brain_pval", "lme.apoe.brain_pval",
        "lme.apoe_pval", "lme.apoe_beta")

mega.apoe = prepare_df2(df.mega.apoe, p.correct, p.direction, vars_sign=c())

save(mega.apoe,
       file = file.path(wd,"mega","apoe", "df.mega.apoe.rda"))
```

## apoe plots

```{r apoe plots}
apoe_plots(wd)    
```

## get predictions and other stuff

```{r}
load(file.path(wd,"mega","apoe", "df.mega.apoe.rda"))
dt.predictions = wrapper_get_predictions_derivative.apoe(wd) 
dt.predictions = data.table::rbindlist(dt.predictions)
mega.apoe = left_join(mega.apoe, dt.predictions)    
mega.apoe = wrapper_get_estimates_apoe(mega.apoe) 
dt.estimates = wrapper_draw_smooth_apoe(mega.apoe)
dt.estimates = data.table::rbindlist(dt.estimates)
mega.apoe = left_join(mega.apoe, dt.estimates)       
save(mega.apoe, file =file.path(wd,"mega","apoe", "df.mega.apoe.rda"))
```

### compare apoe - main (SI)

```{r}
load(file =file.path(wd,"mega", "df.mega.rda"))
load(file.path(wd,"mega","apoe", "df.mega.apoe.rda"))

features = mega$features

grot = list()
for (i in 1:length(features)) {
  print(i)
  X = mega %>% filter(features == features[[i]]) %>% .$datamainsmooth %>% .[[1]]
  ss = mega.apoe %>% filter(features == features[[i]]) %>%  .$data.predict.main.apoeO %>% .[[1]] %>% filter(is.na(apoe_statusO))
  
  X$estimate.apoe0 = approx(ss$delta_brain, ss$.estimate, xout = X$delta_brain)$y
  grot[[i]]  = 
    X %>% 
    rowwise() %>% 
    mutate(is.equal = between(estimate.apoe0, .estimate - 1.96*.se, .estimate + 1.96*.se), 
           features = features[[i]]) %>% 
    group_by(features) %>% 
    nest()
}

df.grot = data.table::rbindlist(grot)

x = df.grot %>% unnest() 
table(x$is.equal, useNA = "always")
xx = x %>% filter(!is.equal == 1)
#100 - 100*66/(116459+66)

sig.features = mega %>% filter(gamm.w.pval.bootstrap.fdr  < .05) %>% .$features
sig.features.apoe0 = mega.apoe %>% filter(gam.main.w.apoe0p.bootstrap < .05) %>% .$features
setdiff(sig.features, sig.features.apoe0)
```

### compare apoe - interaction (SI)

```{r}
load(file =file.path(wd,"mega", "df.mega.rda"))
load(file.path(wd,"mega","apoe", "df.mega.apoe.rda"))

features = mega$features

grot = list()
for (i in 1:length(features)) {
  print(i)
  X = mega %>% filter(features == features[[i]]) %>% .$datainteractionsmooth %>% .[[1]]
  ss = mega.apoe %>% filter(features == features[[i]]) %>%  .$data.predict.int.apoeO %>% .[[1]] %>% filter(apoe_statusO == 0)
  
  ages = c(40,50,60,70,80)
  X = X %>% filter(xage %in% ages) %>% 
    group_by(xage) %>% 
    nest()
  ss = ss %>% filter(xage %in% ages) %>% 
    group_by(xage) %>% 
    nest() %>% 
    rename(data.apoe = data)
  X = left_join(X, ss)
  X = 
    X %>% 
    mutate(data = 
             map2(data, data.apoe, ~ mutate(.x, estimate.apoe = approx(.y$delta_brain, .y$estimate, xout = .x$delta_brain)$y)))
 
  X = X %>% select(xage, data) %>% unnest()
  
  grot[[i]]  = 
    X %>% 
    rowwise() %>% 
    mutate(is.equal = between(estimate.apoe, estimate - 1.96*se, estimate + 1.96*se), 
           features = features[[i]]) %>% 
    group_by(features) %>% 
    nest()
}

df.grot = data.table::rbindlist(grot)

x = df.grot %>% unnest() 
table(x$is.equal, useNA = "always")
xx = x %>% filter(!is.equal == 1)
#100*1363/(16067+1363)

sig.features = mega %>% filter(gamm.int.w.pval.bootstrap.fdr  < .05) %>% .$features
sig.features.apoe0 = mega.apoe %>% filter(gam.int.w.apoe0p.bootstrap < .05) %>% .$features
setdiff(sig.features, sig.features.apoe0)
```

# 5 Simulation.

```{r simulation}
outdir = file.path(wd, "mega", "report")
outdir.mega.simulation = file.path(outdir, "simulation")
set.seed(1234)

mini.grid = 
  expand.grid(n = 3700, 
              x1 = seq(-1.2, -.2, by= .4), 
              sd1 = seq(.6,1.8, by = .6), 
              skew1 = -5,
              sd2 = .8, 
              r = .3,
              ysd = .8)


out.mg = out.dataset.mg = df.tidy.mg = smt.model.mg = smt.aging.mg = list()
for(i in 1:nrow(mini.grid)) {
  print(i)
  out.mg[[i]] = replicate(1000, model_twoModel_discussion(mini.grid$n[[i]], mini.grid$x1[[i]], mini.grid$sd1[[i]], mini.grid$skew1[[i]], mini.grid$sd2[[i]], mini.grid$r[[i]], mini.grid$ysd[[i]]), simplify = F)
  out.dataset.mg[[i]] = twoModel_distribution_example(100000, mini.grid$x1[[i]], mini.grid$sd1[[i]], mini.grid$skew1[[i]], mini.grid$sd2[[i]], mini.grid$r[[i]], mini.grid$ysd[[i]])
  df.tidy.mg[[i]] = tidy_stats(out.mg[[i]])
  smt.model.mg[[i]] = get_mean_smooths(out.mg[[i]], smooth = "model")
  smt.aging.mg[[i]] = get_mean_smooths(out.mg[[i]], smooth = "aging")
  
}


mini.grid= mini.grid %>% 
  mutate(out = out.mg, 
         out.dataset = out.dataset.mg, 
         df.tidy = df.tidy.mg, 
         smt.model = smt.model.mg, 
         smt.aging = smt.aging.mg)

save(mini.grid, 
     file =  file.path(outdir.mega.simulation, "modelling.rda"))


skew.grid = 
  expand.grid(n = 3700, 
              x1 = -.8, 
              sd_target = .748, # approximate sd for main simulation
              skew1 = c(-2,-5, -20),
              sd2 = .8, 
              r = .3,
              ysd = .8)



out.mg = out.dataset.mg = df.tidy.mg = smt.model.mg = smt.aging.mg = list()
for(i in 1:nrow(skew.grid)) {
  print(i)
  skew.grid$sd1[i] = adjust_omega(skew.grid$skew1[i], skew.grid$sd_target[i])
  out.mg[[i]] = replicate(1000, model_twoModel_discussion(skew.grid$n[[i]], skew.grid$x1[[i]], skew.grid$sd1[[i]], skew.grid$skew1[[i]], skew.grid$sd2[[i]], skew.grid$r[[i]], skew.grid$ysd[[i]]), simplify = F)
  out.dataset.mg[[i]] = twoModel_distribution_example(100000, skew.grid$x1[[i]], skew.grid$sd1[[i]], skew.grid$skew1[[i]], skew.grid$sd2[[i]], skew.grid$r[[i]], skew.grid$ysd[[i]])
  df.tidy.mg[[i]] = tidy_stats(out.mg[[i]])
  smt.model.mg[[i]] = get_mean_smooths(out.mg[[i]], smooth = "model")
  smt.aging.mg[[i]] = get_mean_smooths(out.mg[[i]], smooth = "aging")
  
}


skew.grid= skew.grid %>% 
  mutate(out = out.mg, 
         out.dataset = out.dataset.mg, 
         df.tidy = df.tidy.mg, 
         smt.model = smt.model.mg, 
         smt.aging = smt.aging.mg)

save(skew.grid, file = file.path(outdir.mega.simulation, "skewgrid.rda"))
```

