rm---
title: "7_FIGURES"
author: "MS"
date: '2023-11-13'
output: html_document
---

#Load lib
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gratia)
library(here)
library(tidyverse)
library(mgcv)
library(viridis)
library(patchwork)
library(kableExtra)
library(ggseg)
library(ggsegDesterieux)
library(extrafont)
library(ggridges)
options(bitmapType='cairo')

wd = here("data_memory_long/all")
outdir = file.path(wd, "mega", "report")
load(file.path(wd, "desterieux.Rda"))
source(here("scripts/memory_long_scripts/7_figures_functions.R"))
```

# FIGURES
#1.MEGA (change) 
## 1.1Main
```{r mega change main, echo=FALSE, warning=F}
outdir.mega.main = file.path(outdir, "main")
load(file.path(wd,"mega", "df.mega.rda"))

df <- mega %>% 
    ungroup() 


(m2 <- brain_plots(df, "gamm.w.beta.m5.0.weighted"))
ggsave(file.path(outdir.mega.main, "main.unthreshold.cortical.png"), plot = m2$g1)
ggsave(file.path(outdir.mega.main, "main.unthreshold.aseg.png"), plot = m2$g2)


(m1 <- wrapper_brain_plots(df, "gamm.w.beta.m5.0.weighted"))
ggsave(file.path(outdir.mega.main, "main.threshold.cortical.png"), plot = m1$g1)
ggsave(file.path(outdir.mega.main, "main.threshold.aseg.png"), plot = m1$g2)


dat = mega %>% 
  filter(gamm.w.pval.bootstrap < .05) %>% 
  select(features, 
         gamm.w.beta.m5.0.weighted, 
         gamm.w.beta.m2.0.weighted, 
         gamm.w.pval.bootstrap.log, 
         gamm.w.pval.bootstrap.log.fdr) %>% 
  arrange(desc(gamm.w.pval.bootstrap.log))
(kat = dat %>% kable(digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))
writeLines(kat, file.path(outdir.mega.main, "main.table.sig.html"))

#load(file.path(wd, "merged_reliability.rda"))
#xx = df.merged$data[[1]]$delta_brain
#quantile(xx, c(0.025, 0.975)) 
gs = wrapper_plot_trajectories_main(mega, outdir.mega.main)
```

## 1.2 Age interaction
```{r mega change interaction I , echo=FALSE, warning=F}
outdir.mega.int = file.path(outdir, "int")
load(file.path(wd,"mega", "df.mega.rda"))

df <- mega %>% 
    ungroup() 
 
(m2 <- brain_plots(df, "gamm.int.w.pval.bootstrap.log"))
ggsave(file.path(outdir.mega.int, "int.unthreshold.cortical.png"), plot = m2$g1)
ggsave(file.path(outdir.mega.int, "int.unthreshold.aseg.png"), plot = m2$g2)

(m1 <-  wrapper_brain_plots(df, "gamm.int.w.pval.bootstrap.log", mode = "int"))
ggsave(file.path(outdir.mega.int, "int.threshold.cortical.png"), plot = m1$g1)
ggsave(file.path(outdir.mega.int, "int.threshold.aseg.png"), plot = m1$g2)

dat = mega %>% 
  filter(gamm.int.w.pval.bootstrap.log > 1.3) %>% 
  select(features, gamm.int.w.pval.bootstrap.log, gamm.int.w.pval.bootstrap.log.fdr) %>% 
  arrange(desc(gamm.int.w.pval.bootstrap.log.fdr))
(kat = dat %>% kable(digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))
writeLines(kat, file.path(outdir.mega.int, "int.table.sig.html"))


gs = wrapper_plot_trajectories_int(df)
```

# 1.3 Hemi diff
```{r}
gs = wrapper_hemi_plot_main(wd, outdir)
gs.errorbar = wrapper_hemi_plot_int(wd, outdir)  
```


# 2.REGIONAL VS GLOBAL EFFECTS 
```{r clustering pca and corr matrix, echo=FALSE, warning=F}
outdir = file.path(wd, "mega", "report")
gs.cor = wrapper_density_tile_plot(wd, outdir)
gs.cc = wrapper_consensus_cluster_plot(wd, outdir)
gs.pca = wrapper_plot_pca(wd, outdir)

load(file.path(wd,"mega","consensus_cluster","mod.results.rda"))


df.cl1 = df.cl1 %>% 
  mutate(type = "cl")
df.pca = df.pca %>% 
  mutate(type = "pca")
df.basic = df.basic %>% 
  mutate(type = "basic")
df.clall = df.clall %>% 
  mutate(type = "all") %>% 
  select(cl, pval, pboot, type)
df.clall$cl = df.basic$cl

df= reduce(list(df.cl1, 
            df.pca, 
            df.basic, 
            df.clall), 
       rbind)

dt.cl11 = dt.cl11 %>% 
  mutate(type = "cl")
dt.cl11$cl = df.cl1$cl

dt.pca1 = dt.pca1 %>% 
  mutate(type = "pca")
dt.pca1$cl = df.pca$cl

dt.basic1 = dt.basic1 %>% 
  mutate(type = "basic")
dt.basic1$cl = df.basic$cl

dt.clall1 = dt.clall1 %>% 
  mutate(type = "all") 
dt.clall1$cl = dt.basic1$cl
  
dt= reduce(list(dt.cl11, 
            dt.pca1, 
            dt.basic1, 
            dt.clall1), 
       rbind)

dat = left_join(df, dt)
(kat = dat %>% kable(digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))
writeLines(kat,  file.path(outdir, "dimensionality", "table.mod.html"))


```  


#3.APOE 

### 3.1 Apoe on memory
```{r}
outdir.mega.mem = file.path(outdir, "memory_apoe")
gs = wrapper_memory_plot_main(wd, outdir) 
gs1 = wrapper_memory_plot_age(wd, outdir)
gsall = gs+gs1
ggsave(file.path(outdir.mega.mem, "combinedplots.png"),
                 plot = gsall, 
                 height = 7,
                 width = 14)
```


### 3.2 Apoe on brain
#### 3.2.1 Main effect
```{r}
#table
 outdir.mega.brain= file.path(outdir, "brain_apoe")
  load( file.path(wd,"mega","apoe", "df.mega.apoe.brain.rda"))
  dat = mega.apoe.brain %>% 
    filter(lme.apoe.brain_pval  < 0.05) %>% 
    select(features, 
           lme.apoe.brain_beta,
           lme.apoe.brain_pval.log,
           lme.apoe.brain_pval.log.fdr) %>% 
    arrange(lme.apoe.brain_beta)
  (kat = dat %>% kable(digits = 3) %>% 
      kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat, file.path(outdir.mega.brain, "table.sig.main.html"))
  
gs = wrapper_plot_apoe_brain_main(wd, outdir) 
gs1 = wrapper_plot_apoe_brain_main_across(wd, outdir)  
gsall = gs + gs1 + plot_layout(widths = c(5, 1.2))
ggsave(file.path(outdir.mega.brain, "apoe_brain_combined.png"), 
       plot = gsall, width = 16, height = 7)  
  
```

#### 3.2.2 Age effects
```{r}
# table
 outdir.mega.brain= file.path(outdir, "brain_apoe")
  load( file.path(wd,"mega","apoe", "df.mega.apoe.brain.rda"))
  dat = mega.apoe.brain %>% 
    filter(gam.apoe.brain_agep.boot  < 0.05) %>% 
    select(features, 
           gam.apoe.brain_ageF,
           gam.apoe.brain_agep.boot.log,
           gam.apoe.brain_agep.boot.log.fdr) %>% 
    arrange(gam.apoe.brain_agep.boot.log)
  (kat = dat %>% kable(digits = 3) %>% 
      kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat, file.path(outdir.mega.brain, "table.sig.age.html"))
  
 
  gsall = wrapper_plot_apoe_brain_age(wd, outdir) 
```


### 3.3 Apoen on brain - memory associations
```{r}
load( file.path(wd,"mega","apoe", "df.mega.apoe.rda"))
outdir.mega.apoe= file.path(outdir, "apoe")
dat  = 
  mega.apoe %>% 
  select(features, 
         gam.main.w.apoeOF, 
         gam.main.w.apoeOp, 
         gam.main.w.apoeOp.bootstrap.log, 
         gam.main.w.apoeOp.bootstrap.log.fdr) %>% 
  filter(gam.main.w.apoeOp.bootstrap.log > 1.3) %>% 
  arrange(gam.main.w.apoeOp.bootstrap.log)
  (kat = dat %>% kable(digits = 3) %>% 
      kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat, file.path(outdir.mega.apoe, "table.sig.main.html"))                          

  # table interaction
  dat1  = 
  mega.apoe %>% 
  select(features, 
         gam.int.w.apoeOF, 
         gam.int.w.apoeOp, 
         gam.int.w.apoeOp.bootstrap.log, 
         gam.int.w.apoeOp.bootstrap.log.fdr) %>% 
  filter(gam.int.w.apoeOp.bootstrap.log > 1.3) %>% 
  arrange(gam.int.w.apoeOp.bootstrap.log)
  (kat1 = dat1 %>% kable(digits = 3) %>% 
      kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat1, file.path(outdir.mega.apoe, "table.sig.int.html"))                          


### Plot  basic
  gs = wrapper_plot_apoe_main(wd, outdir)
  gs1 = wrapper_plot_apoe_int(wd, outdir)

  gsall = gs + gs1 + plot_layout(widths = c(1, 3))
  ggsave(file.path(outdir.mega.apoe, "apoe_hipp_combined.png"), 
       plot = gsall, width = 14, height = 8)  
```

# 4 SIMULATIONS
```{r simulation}
outdir = file.path(wd, "mega", "report")
outdir.mega.simulation = file.path(outdir, "simulation")
gs1.nonlinear = wrapper_simulation_nonlinear(outdir.mega.simulation) # linear to non-linear
gs2.meanchange = wrapper_simulation_meanchange(outdir.mega.simulation) # mean change
gs3.dispersion = wrapper_simulation_dispersion(outdir.mega.simulation )
gs4.skew = wrapper_simulation_skew(outdir.mega.simulation)
gs = gs1.nonlinear[[1]] + gs2.meanchange[[1]] + gs3.dispersion[[1]] + gs4.skew[[1]]+ plot_layout(ncol = 4)
ggsave(file.path(outdir.mega.simulation, "simulation.all.png"), 
         plot = gs, 
         width = 24, 
         height = 8)
```

## provitional code to test mean effects
```{r}
#  out.dataset.X = 
#   rbind(
#   df$out.dataset[[1]] %>% 
#   mutate(g = "g") %>% 
#   pivot_longer(-g, 
#                names_to = "G", 
#                values_to = "X") %>% 
#   filter(G %in% c("X")) %>% 
#     mutate(skew = "low"), 
#   df$out.dataset[[2]] %>% 
#   mutate(g = "g") %>% 
#   pivot_longer(-g, 
#                names_to = "G", 
#                values_to = "X") %>% 
#   filter(G %in% c("X")) %>% 
#     mutate(skew = "mid"), 
#   df$out.dataset[[3]] %>% 
#   mutate(g = "g") %>% 
#   pivot_longer(-g, 
#                names_to = "G", 
#                values_to = "X") %>% 
#   filter(G %in% c("X")) %>% 
#     mutate(skew = "high")
# )
#  
# out.dt =  out.dataset.X %>% 
#    group_by(skew) %>% 
#    summarise(mean(X), 
#              sd(X))
# 
# print((out.dt$`mean(X)`[3] - out.dt$`mean(X)`[2])/out.dt$`sd(X)`[1])
```


#TABLES
# 1. Main demographics
```{r}
outdir.mega.tables = file.path(outdir, "tables")
load(file.path(wd, "merged_reliability.rda"))
df = 
  df.merged %>% 
  filter(features == "Left-Hippocampus") %>%  
  unnest()


dat = 
    df %>% 
    group_by(dataset) %>% 
    summarise(Subs = n(),
              sex = sum(sex), 
              apoeNA = sum(is.na(apoe_status)),
              apoe0 = sum(apoe_status == 0, na.rm = T),
              apoe1 = sum(apoe_status == 1, na.rm = T),
              xage.mean = mean(xage),
              xage.sd = sd(xage) , 
              xage.min =min(xage), 
              xage.max = max(xage),
              Obs.mem = sum(total_tp_memory), 
              Obs.mem.mean = mean(total_tp_memory), 
              Obs.mem.sd = sd(total_tp_memory), 
              Obs.mem.min = min(total_tp_memory), 
              Obs.mem.max = max(total_tp_memory), 
              Obs.brain = sum(n), 
              Obs.brain.mean = mean(n), 
              Obs.brain.sd = sd(n), 
              Obs.brain.min = min(n), 
              Obs.brain.max = max(n), 
              time.brain.mean = mean(time), 
              time.brain.sd = sd(time), 
              time.brain.min = min(time), 
              time.brain.max = max(time), 
              time.mem.mean = mean(total_followup_memory), 
              time.mem.sd = sd(total_followup_memory), 
              time.mem.min = min(total_followup_memory), 
              time.mem.max = max(total_followup_memory), 
              .groups = 'drop') %>% 
    bind_rows(
      df %>% 
        ungroup() %>% 
        summarise(dataset = "all", 
                  Subs = n(),
                  sex = sum(sex), 
                  apoeNA = sum(is.na(apoe_status)),
                  apoe0 = sum(apoe_status == 0, na.rm = T),
                  apoe1 = sum(apoe_status == 1, na.rm = T),
                  xage.mean = mean(xage),
                  xage.sd = sd(xage) , 
                  xage.min =min(xage), 
                  xage.max = max(xage),
                  Obs.mem = sum(total_tp_memory), 
                  Obs.mem.mean = mean(total_tp_memory), 
                  Obs.mem.sd = sd(total_tp_memory), 
                  Obs.mem.min = min(total_tp_memory), 
                  Obs.mem.max = max(total_tp_memory), 
                  Obs.brain = sum(n), 
                  Obs.brain.mean = mean(n), 
                  Obs.brain.sd = sd(n), 
                  Obs.brain.min = min(n), 
                  Obs.brain.max = max(n), 
                  time.brain.mean = mean(time), 
                  time.brain.sd = sd(time), 
                  time.brain.min = min(time), 
                  time.brain.max = max(time), 
                  time.mem.mean = mean(total_followup_memory), 
                  time.mem.sd = sd(total_followup_memory), 
                  time.mem.min = min(total_followup_memory), 
                  time.mem.max = max(total_followup_memory)))


kat = kable_2_table(dat, bold = "last", digit = 3)
writeLines(kat, file.path(outdir.mega.tables, "main.table.html"))
```


# S Table1. main  demographics. initial  mri sample
```{r}
outdir.mega.tables = file.path(outdir, "tables")
load( "/ess/p274/cluster/projects/p039_image_brain/data_normative_long/df_mri/all/df.all.filt.Rda")
load(file.path(wd,"mega","apoe", "output","APOE_merged.rda"))
merged_apoe_sub = merged_apoe %>% select(rid, apoe4, apoe_status)
df = left_join(df.merge, merged_apoe_sub)
 

df = 
  df %>% 
  group_by(rid) %>% 
  mutate(Obs = n()) %>% 
  filter(!dataset %in% c("ucam", "wayne"))

df.sub = df %>% 
  group_by(rid) %>% 
    summarise(n = first(Obs),
              sex = first(sex), 
              apoe_status = first(apoe_status, na_rm = T),
              xage = mean(age),
              time = max(age) - min(age), 
              dataset = first(dataset)) 

dat = 
    df.sub %>% 
    group_by(dataset) %>% 
    summarise(Subs = n(),
              sex = sum(sex), 
              apoeNA = sum(is.na(apoe_status)),
              apoe0 = sum(apoe_status == 0, na.rm = T),
              apoe1 = sum(apoe_status == 1, na.rm = T),
              xage.mean = mean(xage),
              xage.sd = sd(xage) , 
              xage.min =min(xage), 
              xage.max = max(xage),
              Obs.brain = sum(n), 
              Obs.brain.mean = mean(n), 
              Obs.brain.sd = sd(n), 
              Obs.brain.min = min(n), 
              Obs.brain.max = max(n), 
              time.brain.mean = mean(time), 
              time.brain.sd = sd(time), 
              time.brain.min = min(time), 
              time.brain.max = max(time), 
              .groups = 'drop') %>% 
    bind_rows(
      df.sub %>% 
        ungroup() %>% 
        summarise(dataset = "all", 
                  Subs = n(),
                  sex = sum(sex), 
                  apoeNA = sum(is.na(apoe_status)),
                  apoe0 = sum(apoe_status == 0, na.rm = T),
                  apoe1 = sum(apoe_status == 1, na.rm = T),
                  xage.mean = mean(xage),
                  xage.sd = sd(xage) , 
                  xage.min =min(xage), 
                  xage.max = max(xage),
                  Obs.brain = sum(n), 
                  Obs.brain.mean = mean(n), 
                  Obs.brain.sd = sd(n), 
                  Obs.brain.min = min(n), 
                  Obs.brain.max = max(n), 
                  time.brain.mean = mean(time), 
                  time.brain.sd = sd(time), 
                  time.brain.min = min(time), 
                  time.brain.max = max(time)))


kat = kable_2_table(dat, bold = "last", digit = 3)
writeLines(kat, file.path(outdir.mega.tables, "S1table_demographics_initial_mri_sample.html"))
kat = kable_2_table(dat, bold = "last", digit = 1)
writeLines(kat, file.path(outdir.mega.tables, "S1table_demographics_initial_mri_sample2.html"))
```

# S Table2. main  demographics. initial  cognition sample
```{r}
outdir.mega.tables = file.path(outdir, "tables")
load(file.path(wd, "df.memory_merged_full.rda"))
load(file.path(wd,"mega","apoe", "output","APOE_merged.rda"))
merged_apoe_sub = merged_apoe %>% select(rid, apoe4, apoe_status)
df = left_join(memory.merge.full, merged_apoe_sub)
 

df = 
  df %>% 
  group_by(rid) %>% 
  mutate(Obs = n()) %>% 
  filter(!dataset == "ucam")

tardf.sub = df %>% 
  group_by(rid) %>% 
    summarise(n = first(Obs),
              sex = first(sex), 
              apoe_status = first(apoe_status, na_rm = T),
              xage = mean(age),
              time = max(age) - min(age), 
              dataset = first(dataset)) 

dat = 
    df.sub %>% 
    group_by(dataset) %>% 
    summarise(Subs = n(),
              sex = sum(sex), 
              apoeNA = sum(is.na(apoe_status)),
              apoe0 = sum(apoe_status == 0, na.rm = T),
              apoe1 = sum(apoe_status == 1, na.rm = T),
              xage.mean = mean(xage),
              xage.sd = sd(xage) , 
              xage.min =min(xage), 
              xage.max = max(xage),
              Obs.mem = sum(n), 
              Obs.mem.mean = mean(n), 
              Obs.mem.sd = sd(n), 
              Obs.mem.min = min(n), 
              Obs.mem.max = max(n), 
              time.mem.mean = mean(time), 
              time.mem.sd = sd(time), 
              time.mem.min = min(time), 
              time.mem.max = max(time), 
              .groups = 'drop') %>% 
    bind_rows(
      df.sub %>% 
        ungroup() %>% 
        summarise(dataset = "all", 
                  Subs = n(),
                  sex = sum(sex), 
                  apoeNA = sum(is.na(apoe_status)),
                  apoe0 = sum(apoe_status == 0, na.rm = T),
                  apoe1 = sum(apoe_status == 1, na.rm = T),
                  xage.mean = mean(xage),
                  xage.sd = sd(xage) , 
                  xage.min =min(xage), 
                  xage.max = max(xage),
                  Obs.mem = sum(n), 
                  Obs.mem.mean = mean(n), 
                  Obs.mem.sd = sd(n), 
                  Obs.mem.min = min(n), 
                  Obs.mem.max = max(n), 
                  time.mem.mean = mean(time), 
                  time.mem.sd = sd(time), 
                  time.mem.min = min(time), 
                  time.mem.max = max(time)))


kat = kable_2_table(dat, bold = "last", digit = 3)
writeLines(kat, file.path(outdir.mega.tables, "S2table_demographics_initial_mem_sample.html"))
kat = kable_2_table(dat, bold = "last", digit = 1)
writeLines(kat, file.path(outdir.mega.tables, "S2table_demographics_initial_mem_sample2.html"))
```
# S Table3. Clustering + dimensionality
```{r}
  outdir.mega.tables = file.path(outdir, "tables")
  load(file.path(wd,"mega","consensus_cluster", "cluster_mod.rda"))
  annon <- m3c$mod$realdataresults[[8]]$ordered_annotation
  annon$features = gsub("\\.","-",rownames(annon))
  annon$ccf = as.factor(annon$consensuscluster)
  annon = annon %>% 
    mutate(label = if_else(features == "Left-Inf-Lat-Vent", "Left-Lateral-Ventricle", features))
  
  #
  load(file.path(wd,"mega","PCA" ,"pca_result.rda"))
  df.loading = data.frame(features = rownames(pca_result$rotation), 
                          loading1 = -pca_result$rotation[,1], 
                          loading2 = -pca_result$rotation[,2], 
                          loading3 = -pca_result$rotation[,3], 
                          loading4 = -pca_result$rotation[,4])
  
  
  
  df.changenames = retrieve_orig2plot_names()
  df.loading = 
    df.loading %>% arrange(desc(loading1)) %>% 
    mutate(id = 1:nrow(.) %>% as.factor())
  
  dat = inner_join(df.loading, annon)
  (kat = dat %>% kable(digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat, file.path(outdir.mega.tables , "S3table_dim.loads.html"))
```


# S Table4. regional contributions
```{r}
  outdir.mega.tables = file.path(outdir, "tables")
  

df.basic =retrieve_pboot_cc(wd, type = "basic") %>% 
  mutate(type = "basic")
df.pca =retrieve_pboot_cc(wd, type = "pca") %>% 
  mutate(type = "pca")
df.cl1 =retrieve_pboot_cc(wd, type = "cl1")%>% 
  mutate(type = "cl1")


df.cluster = 
  reduce(list(df.basic, 
            df.pca, 
            df.cl1), 
       rbind) %>% 
  dplyr::select(-pval) %>% 
  pivot_wider(names_from = "type", 
              values_from = "pboot")

  
  (kat = dat %>% kable(digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))
  writeLines(kat, file.path(outdir.mega.tables , "S4cluster.results.html"))
```

# SUPPLEMENTARY FIGURES
# S Fig.1 age distribution
```{r}
outdir.mega.SFigs = file.path(outdir, "SFigs")
load(file.path(wd, "merged_reliability.rda"))
df = 
  df.merged %>% 
  filter(features == "Left-Hippocampus") %>%  
  unnest()

gs = 
ggplot(df, aes(x = age, y = dataset, fill = 0.5 - abs(0.5 - after_stat(ecdf)), group = dataset)) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE, bandwidth = 2,
                      jittered_points = TRUE, position = "raincloud", point_alpha = .6, point_fill = "#1F77B4", point_color = "black", point_shape = 21) +
  scale_fill_viridis_c(name = "", direction = -1) +
 # scale_discrete_manual(aesthetics = "point_shape", values = c(18)) +
  theme_ridges() +
  theme(legend.position = "none") + 
  labs(y = "Age") +
  theme_minimal(base_size = 16) +
  theme(
    text = element_text(family = "Arial", face =  "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", size = 1),
    axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
    axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
    axis.title.x = element_text(size = 20, family = "Arial", color = "black"),
    axis.title.y = element_blank(),
    plot.title = element_text(size = 24, face = "bold", family = "Arial", hjust = 0.5),
    legend.position = "none"  # Hide legend in individual plot
  ) 

ggsave(file.path(outdir.mega.SFigs, "SFig1.AgedistributionSample.jpeg"), plot = gs, width = 8, height = 10)
```

# S Fig.3. M3c extra figures
```{r}
outdir.mega.SFigs = file.path(outdir, "SFigs")
load(file.path(wd,"mega","consensus_cluster", "cluster_mod.rda"))
gs1 = m3c$mod$plots[[1]]
gs2 = m3c$mod$plots[[2]]
gs3 = m3c$mod$plots[[3]]
gs4 = m3c$mod$plots[[4]]


gs1 = 
gs1 + labs(x = "Consensus Index", 
           colour = "K") + 
  theme_classic() +
  theme(legend.position = "right", 
        text = element_text(family = "Arial", face =  "bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 1),
        axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
        axis.title = element_text(size = 20, family = "Arial", color = "black"),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        plot.title = element_blank()) + 
   guides(color = guide_legend(keywidth = 3, keyheight = 1))

gs2 = 
  gs2+ 
  theme_classic() +
  theme(legend.position = "bottom", 
        text = element_text(family = "Arial", face =  "bold"),
        axis.line = element_line(color = "black", size = 1),
        axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
        axis.title = element_text(size = 20, family = "Arial", color = "black"),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        plot.title = element_blank())

gs3 = 
  gs3+ 
  theme_classic() +
  theme(legend.position = "bottom", 
        text = element_text(family = "Arial", face =  "bold"),
        axis.line = element_line(color = "black", size = 1),
        axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
        axis.title = element_text(size = 20, family = "Arial", color = "black"),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        plot.title = element_blank())



gs4 = 
  gs4+ 
  theme_classic() +
  theme(legend.position = "bottom", 
        text = element_text(family = "Arial", face =  "bold"),
        axis.line = element_line(color = "black", size = 1),
        axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
        axis.title = element_text(size = 20, family = "Arial", color = "black"),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        plot.title = element_blank())


gs = gs1 / (gs2 + gs3 + gs4)
ggsave(file.path(outdir.mega.SFigs, "SFig3.cluster.jpeg"), plot = gs, width = 9, height = 12)

```


# DEPRECATED
#### apoe table
```{r}
load(file.path(wd,"mega","apoe", "output" ,"APOE_merged.rda"))

#TABLE
summary_table <- merged_apoe %>%
  filter(!dataset == "ucam") %>% 
  group_by(dataset) %>%
  summarize(
    APOE_data = sum(!is.na(apoe_status)),  # Count non-NA values
    APOE_na = sum(is.na(apoe_status)),     # Count NA values
    carriers = sum(apoe_status == 1, na.rm = TRUE),  # Count where apoe_status is 1 (carriers)
    non_carriers = sum(apoe_status == 0, na.rm = TRUE)  # Count where apoe_status is 0 (non-carriers)
  ) %>%
  ungroup()  # Remove the grouping

# Add a total summary row
total_summary <- summary_table %>%
  filter(!dataset == "ucam") %>% 
  summarize(
    dataset = "TOTAL",
    APOE_data = sum(APOE_data),
    APOE_na = sum(APOE_na),
    carriers = sum(carriers),
    non_carriers = sum(non_carriers)
  )

# Combine dataset summaries with the total summary
final_table <- bind_rows(summary_table, total_summary)
(kabapoe = 
  kable(final_table, digits = 3, caption = "apoe overview") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F))

save(final_table,
     kabapoe,
     file = file.path(wd,"mega","apoe", "output", "APOE_overview.rda"))

```


# S5. memory scores
```{r}
load(file.path(wd, "df.memory_merged_full.rda"))
memory.merge.full = 
  memory.merge.full %>% 
  filter(!dataset == "ucam")
gs = ggplot(memory.merge.full, aes(age, PCA, color= dataset)) + 
  geom_point() + 
  geom_smooth(color = "black", size = 2) +
  theme_classic()  + 
  facet_wrap(.~dataset, ncol = 3, scales = "free_y")  +
  theme(legend.position = "none", 
        text = element_text(family = "Arial", face =  "bold"),
        axis.line = element_line(color = "black", size = 1),
        axis.text.x = element_text(size = 18, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 18, family = "Arial", color = "black"),
        axis.title = element_text(size = 20, family = "Arial", color = "black"),
        strip.text = element_text(size = 18, family = "Arial", color = "black"),
        strip.background = element_blank(), 
        plot.title = element_blank())
ggsave(file.path(outdir.mega.SFigs, "SFig5.memory.jpeg"), plot = gs, width = 10, height = 10)

```